$K$-corrections and filter transformations \\ in the ultraviolet, optical, and near infrared

### 2006-06-07 docs/paper/kcorrect-aj-v1/ms.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@553 75dff457-8751-0410-ab1a-cd701654ce6b

\documentclass[10pt,preprint]{aastex}
\usepackage{amsmath}
\usepackage{epsfig}

@@

\newcommand{\XX}{X}
\newcommand{\XXt}{\tilde{X}}
\newcommand{\XXh}{\hat{X}}
\newcommand{\Xkn}{X_{kn}}
\newcommand{\Xknh}{\hat{X}_{kn}}
\renewcommand{\AA}{A}
\newcommand{\AAt}{\tilde{A}}
\newcommand{\WW}{W}
\newcommand{\WWt}{\tilde{W}}
\newcommand{\Wki}{W_{ki}}
\newcommand{\HH}{H}
\newcommand{\HHt}{\tilde{H}}
\newcommand{\Hin}{H_{in}}
\newcommand{\BB}{B}
\newcommand{\BBt}{\tilde{B}}
\newcommand{\MM}{M}
\newcommand{\MMt}{\tilde{M}}
\newcommand{\Aki}{A_{ki}}
\newcommand{\Bij}{B_{ij}}
\newcommand{\Mjn}{M_{jn}}
\newcommand{\skn}{\sigma_{kn}}

@@

\title{$K$-corrections and filter transformations \\
from the restframe
ultraviolet to the near infrared}

@@

\begin{abstract}
We present a method of developing model-based spectroscopic templates
to explain observed galaxy fluxes. Such templates are useful for
calculating $K$-corrections and converting among observations of
galaxies at various wavelengths. In principle, we could apply this
method to galaxy data at any redshift and in any set of
wavelengths. The implementation we present here is suitable for
ultraviolet, optical, and near-infrared observations in the redshift
range $0 < z < 1.5$. Since we base the templates on stellar population
synthesis results, the results are intepretable in terms of
approximate stellar masses and star-formation histories. We present
templates fit with this method to data from GALEX, Sloan Digital Sky
Survey spectroscopy and photometry, the Two-Micron All Sky Survey, the
Deep Extragalactic Evolutionary Probe and Great Observatories Origins
Deep Survey.  In addition, we present software for using such data to
estimate $K$-corrections, stellar masses, and recent star-formation
rates. This software is suitable for analyzing many other data sets as
well.
\end{abstract}

@@

\bibliographystyle{/home/users/mb144/nyu-astro/tex/apj}
\bibliography{/home/users/mb144/nyu-astro/tex/apj-jour,/home/users/mb144/nyu-astro/tex/ccpp}

@@

\subsection{Standard NMF}
The standard non-negative matrix factorization (NMF) problem, as
originally posed by \cite{lee00a}, is to approximate a data
matrix $\XX$ (of size $N_k\times N_n$) as the outer product of two
rank-$N_i$ non-negative matrices $\WW$ and $\HH$. For example, in this
paper, $\XX$ represents the fluxes at $N_n$ different wavelengths for
$N_k$ different galaxies, $\HH$ represents a set of $N_i$ templates to build
each galaxy from, and $\WW$ represents, for each galaxy, the weights
to give each template. This factorization is similar in spirit to
singular value decomposition (SVD) and principal components analysis
(PCA) but crucially, in NMF, $\XX$, $\WW$ and $\HH$ are all
constrained to have non-negative entries. This changes the problem
fundamentally, since no ``cancellation'' of positive and negative
basis functions or coefficients is possible -- all approximation
interactions are strictly additive. The goal, as with SVD or PCA, is
to minmize the squared approximation error (Frobenius norm):
\beqa
\chi^2 &= \|\XX-\WW\HH\|^2\\
%&= \|\XX-\XXh\|^2\\
&= \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
\eeqa
In their seminal papers, \cite{lee00a} show that this approximation
error in non-increasing under the following very simple multiplicative
update rules: 
\beqa
\Wki &\leftarrow \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
\Hin &\leftarrow \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
\eeqa
$[M]_{ab}$ denotes the $(ab)$ element of the matrix valued expression
$M$. $\HH\T$ is the transpose of matrix $\HH$.

@@

Our main algorithmic contribution is to derive a similar update equation
for the elements of $\BB$:
\beq
\Bij \leftarrow \Bij
\frac{[\AA\T\XX\MM\T]_{ij}}{[\AA\T\AA\BB\MM\MM\T]_{ij}}
\eeq
Following the technique outlined in \cite{lee00a}, it can still be shown
that the error is nonincreasing under the application of this update.

{\bf proof??}

@@

\include{kcorrect_tables}

\include{kcorrect_figures}

>>>>

\documentclass[12pt,preprint]{aastex}
\usepackage{amsmath}
\usepackage{epsfig}

@@

\newcommand{\XX}{X}
\newcommand{\XXt}{\tilde{X}}
\newcommand{\XXh}{\hat{X}}
\newcommand{\Xkn}{X_{kn}}
\newcommand{\Xknh}{\hat{X}_{kn}}
\renewcommand{\AA}{A}
\newcommand{\AAt}{\tilde{A}}
\newcommand{\WW}{W}
\newcommand{\WWt}{\tilde{W}}
\newcommand{\Wki}{W_{ki}}
\newcommand{\Wkl}{W_{k\ell}}
\newcommand{\HH}{H}
\newcommand{\HHt}{\tilde{H}}
\newcommand{\Hin}{H_{in}}
\newcommand{\Hln}{H_{\ell{}n}}
\newcommand{\BB}{B}
\newcommand{\BBt}{\tilde{B}}
\newcommand{\MM}{M}
\newcommand{\MMt}{\tilde{M}}
\newcommand{\Aki}{A_{ki}}
\newcommand{\Akl}{A_{k\ell}}
\newcommand{\Bij}{B_{ij}}
\newcommand{\Bls}{B_{\ell{}s}}
\newcommand{\Mjn}{M_{jn}}
\newcommand{\Msn}{M_{sn}}
\newcommand{\skn}{\sigma_{kn}}
\newcommand{\PP}{P}
\newcommand{\Pni}{P_{ni}}
\newcommand{\ZZ}{Z}
\newcommand{\Zki}{Z_{ki}}
\newcommand{\Zkl}{Z_{k\ell}}
\newcommand{\Zij}{Z_{ij}}
\newcommand{\Zkj}{Z_{kj}}
\newcommand{\Zks}{Z_{ks}}
\newcommand{\zz}{z}
\newcommand{\yy}{y}
\newcommand{\yn}{y_n}
\newcommand{\yi}{y_i}
\newcommand{\zn}{z_n}
\newcommand{\zi}{z_i}
\newcommand{\betaij}{\beta_{ij}}
\newcommand{\betalj}{\beta_{\ell{}j}}

@@

\title{$K$-corrections and filter transformations \\
in the ultraviolet, optical, and near infrared}

@@

\begin{abstract}
Template fits to observed galaxy fluxes allow calculation of
$K$-corrections and conversions among observations of galaxies at
various wavelengths. We present a method for creating model-based
template sets given a set of heterogeneous photometric and
spectroscopic galaxy data. Our technique, non-negative matrix
factorization, is akin to principle component analysis (PCA), except
that it is constrained to produce nonnegative templates, it can use a
basis set of models (rather than the delta function basis of PCA), and
it naturally handles uncertainties, missing data, and heterogeneous
data (including broad-band fluxes at various redshifts).  The
particular implementation we present here is suitable for ultraviolet,
optical, and near-infrared observations in the redshift range $0 < z <
1.5$. Since we base our templates on stellar population synthesis
models, the results are intepretable in terms of approximate stellar
masses and star-formation histories. We present templates fit with
this method to data from GALEX, Sloan Digital Sky Survey spectroscopy
and photometry, the Two-Micron All Sky Survey, the Deep Extragalactic
Evolutionary Probe and the Great Observatories Origins Deep Survey.
In addition, we present software for using such data to estimate
$K$-corrections and stellar masses.
\end{abstract}

@@

\begin{thebibliography}{39}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Bell} \& {de~Jong}(2001)}]{bell01b}
{Bell}, E.~F. \& {de~Jong}, R.~S. 2001, \apj, 550, 212

\bibitem[{{Bertin} \& {Arnouts}(1996)}]{bertin96a}
{Bertin}, E. \& {Arnouts}, S. 1996, \aaps, 117, 393

\bibitem[{{Bessell}(1990)}]{bessell90a}
{Bessell}, M.~S. 1990, \pasp, 102, 1181

\bibitem[{Blanton {et~al.}(2003)Blanton, Brinkmann, Csabai, Doi, Eisenstein,
  Fukugita, Gunn, Hogg, \& Schlegel}]{blanton03b}
Blanton, M.~R., Brinkmann, J., Csabai, I., Doi, M., Eisenstein, D.~J.,
  Fukugita, M., Gunn, J.~E., Hogg, D.~W., \& Schlegel, D.~J. 2003, \aj, 125,
  2348

\bibitem[{{Blanton} {et~al.}(2003){Blanton}, {Lin}, {Lupton}, {Maley}, {Young},
  {Zehavi}, \& {Loveday}}]{blanton03a}
{Blanton}, M.~R., {Lin}, H., {Lupton}, R.~H., {Maley}, F.~M., {Young}, N.,
  {Zehavi}, I., \& {Loveday}, J. 2003, \aj, 125, 2276

\bibitem[{{Blanton} {et~al.}(2005)}]{blanton05a}
{Blanton}, M.~R. {et~al.} 2005, \aj, 129, 2562

\bibitem[{Bruzual \& Charlot(2003)}]{bruzual03a}
Bruzual, A.~G. \& Charlot, S. 2003, \mnras, 344, 1000

\bibitem[{{Chabrier}(2003)}]{chabrier03a}
{Chabrier}, G. 2003, \pasp, 115, 763

\bibitem[{{Cohen} {et~al.}(2003){Cohen}, {Wheaton}, \& {Megeath}}]{cohen03a}
{Cohen}, M., {Wheaton}, W.~A., \& {Megeath}, S.~T. 2003, \aj, 126, 1090

\bibitem[{{Connolly} {et~al.}(1995){Connolly}, {Szalay}, {Bershady}, {Kinney},
  \& {Calzetti}}]{connolly95b}
{Connolly}, A.~J., {Szalay}, A.~S., {Bershady}, M.~A., {Kinney}, A.~L., \&
  {Calzetti}, D. 1995, \aj, 110, 1071

\bibitem[{{Davis} {et~al.}(2003){Davis}, {Faber}, {Newman}, {Phillips},
  {Ellis}, {Steidel}, {Conselice}, {Coil}, {Finkbeiner}, {Koo}, {Guhathakurta},
  {Weiner}, {Schiavon}, {Willmer}, {Kaiser}, {Luppino}, {Wirth}, {Connolly},
  {Eisenhardt}, {Cooper}, \& {Gerke}}]{davis03a}
{Davis}, M., {Faber}, S.~M., {Newman}, J., {Phillips}, A.~C., {Ellis}, R.~S.,
  {Steidel}, C.~C., {Conselice}, C., {Coil}, A.~L., {Finkbeiner}, D.~P., {Koo},
  D.~C., {Guhathakurta}, P., {Weiner}, B., {Schiavon}, R., {Willmer}, C.,
  {Kaiser}, N., {Luppino}, G.~A., {Wirth}, G., {Connolly}, A., {Eisenhardt},
  P., {Cooper}, M., \& {Gerke}, B. 2003, in Discoveries and Research Prospects
  from 6- to 10-Meter-Class Telescopes II. Edited by Guhathakurta, Puragra.
  Proceedings of the SPIE, Volume 4834, pp. 161-172 (2003)., 161--172

\bibitem[{Eisenstein {et~al.}(2001)}]{eisenstein01a}
Eisenstein, D.~J. {et~al.} 2001, \aj, 122, 2267

\bibitem[{{Faber} {et~al.}(2003){Faber}, {Phillips}, {Kibrick}, {Alcott},
  {Allen}, {Burrous}, {Cantrall}, {Clarke}, {Coil}, {Cowley}, {Davis}, {Deich},
  {Dietsch}, {Gilmore}, {Harper}, {Hilyard}, {Lewis}, {McVeigh}, {Newman},
  {Osborne}, {Schiavon}, {Stover}, {Tucker}, {Wallace}, {Wei}, {Wirth}, \&
  {Wright}}]{faber03a}
{Faber}, S.~M., {Phillips}, A.~C., {Kibrick}, R.~I., {Alcott}, B., {Allen},
  S.~L., {Burrous}, J., {Cantrall}, T., {Clarke}, D., {Coil}, A.~L., {Cowley},
  D.~J., {Davis}, M., {Deich}, W.~T.~S., {Dietsch}, K., {Gilmore}, D.~K.,
  {Harper}, C.~A., {Hilyard}, D.~F., {Lewis}, J.~P., {McVeigh}, M., {Newman},
  J., {Osborne}, J., {Schiavon}, R., {Stover}, R.~J., {Tucker}, D., {Wallace},
  V., {Wei}, M., {Wirth}, G., \& {Wright}, C.~A. 2003, in Instrument Design and
  Performance for Optical/Infrared Ground-based Telescopes. Edited by Iye,
  Masanori; Moorwood, Alan F. M. Proceedings of the SPIE, Volume 4841, pp.
  1657-1669 (2003)., 1657--1669

\bibitem[{Fukugita {et~al.}(1996)Fukugita, Ichikawa, Gunn, Doi, Shimasaku, \&
  Schneider}]{fukugita96a}
Fukugita, M., Ichikawa, T., Gunn, J.~E., Doi, M., Shimasaku, K., \& Schneider,
  D.~P. 1996, \aj, 111, 1748

\bibitem[{{Giavalisco} {et~al.}(2004)}]{giavalisco04a}
{Giavalisco}, M. {et~al.} 2004, \apjl, 600, L93

\bibitem[{{Hayes}(1985)}]{hayes85a}
{Hayes}, D.~S. 1985, in IAU Symp. 111: Calibration of Fundamental Stellar
  Quantities, Vol. 111, 225--249

\bibitem[{{Hogg}(1999)}]{hogg99a}
{Hogg}, D.~W. 1999, in ASP Conf. Ser. 193: The Hy-Redshift Universe: Galaxy
  Formation and Evolution at High Redshift, 224--+

\bibitem[{{Hogg} {et~al.}(2001){Hogg}, {Finkbeiner}, {Schlegel}, \&
  {Gunn}}]{hogg01a}
{Hogg}, D.~W., {Finkbeiner}, D.~P., {Schlegel}, D.~J., \& {Gunn}, J.~E. 2001,
  \aj, 122, 2129

\bibitem[{Hogg {et~al.}(2002)}]{hogg02a}
Hogg, D.~W. {et~al.} 2002, \aj, 124, 646

\bibitem[{{Jarrett} {et~al.}(2000){Jarrett}, {Chester}, {Cutri}, {Schneider},
  {Skrutskie}, \& {Huchra}}]{jarrett00a}
{Jarrett}, T.~H., {Chester}, T., {Cutri}, R., {Schneider}, S., {Skrutskie}, M.,
  \& {Huchra}, J.~P. 2000, \aj, 119, 2498

\bibitem[{{Kauffmann} {et~al.}(2003)}]{kauffmann03a}
{Kauffmann}, G. {et~al.} 2003, \mnras, 341, 33

\bibitem[{{Kennicutt} {et~al.}(1994){Kennicutt}, {Tamblyn}, \&
  {Congdon}}]{kennicutt94a}
{Kennicutt}, R.~C., {Tamblyn}, P., \& {Congdon}, C.~E. 1994, \apj, 435, 22

\bibitem[{{Kewley} {et~al.}(2001){Kewley}, {Dopita}, {Sutherland}, {Heisler},
  \& {Trevena}}]{kewley01a}
{Kewley}, L.~J., {Dopita}, M.~A., {Sutherland}, R.~S., {Heisler}, C.~A., \&
  {Trevena}, J. 2001, \apj, 556, 121

\bibitem[{Kurucz(1991)}]{kurucz91a}
Kurucz, R.~L. 1991, in Precision Photometry:\ Astrophysics of the Galaxy, L.
  Davis Press, Inc., 27--44

\bibitem[{Lee \& Seung(1999)}]{lee99a}
Lee, D.~D. \& Seung, H.~S. 1999, Nature, 401, 788

\bibitem[{Lee \& Seung(2000)}]{lee00a}
Lee, D.~D. \& Seung, H.~S. 2000, in {NIPS}, 556--562

\bibitem[{Lupton {et~al.}(2001)Lupton, Gunn, {Ivezi{\'c}}, Knapp, Kent, \&
  Yasuda}]{lupton01a}
Lupton, R.~H., Gunn, J.~E., {Ivezi{\'c}}, Z., Knapp, G.~R., Kent, S., \&
  Yasuda, N. 2001, in ASP Conf.\ Ser.\ 238:\ Astronomical Data Analysis
  Software and Systems X, Vol.~10, 269

\bibitem[{Martin {et~al.}(2005)}]{martin05a}
Martin, D.~C. {et~al.} 2005, \apjl, 619, L1

\bibitem[{{Oke} \& {Sandage}(1968)}]{oke68a}
{Oke}, J.~B. \& {Sandage}, A. 1968, \apj, 154, 21

\bibitem[{Pier {et~al.}(2003)Pier, Munn, Hindsley, Hennessy, Kent, Lupton, \&
  {Ivezi{\' c}}}]{pier03a}
Pier, J.~R., Munn, J.~A., Hindsley, R.~B., Hennessy, G.~S., Kent, S.~M.,
  Lupton, R.~H., \& {Ivezi{\' c}}, {\v Z}. 2003, \aj, 125, 1559

\bibitem[{Richards {et~al.}(2002)}]{richards02a}
Richards, G. {et~al.} 2002, \aj, 123, 2945

\bibitem[{{Schneider} {et~al.}(1983){Schneider}, {Gunn}, \&
  {Hoessel}}]{schneider83a}
{Schneider}, D.~P., {Gunn}, J.~E., \& {Hoessel}, J.~G. 1983, \apj, 264, 337

\bibitem[{Sha {et~al.}(2002)Sha, Saul, \& Lee}]{sha02a}
Sha, F., Saul, L., \& Lee, D. 2002, University of Pennsylvania, Technical
  Report MS-CIS-02-19

\bibitem[{Skrutskie {et~al.}(1997)}]{skrutskie97a}
Skrutskie, M.~F. {et~al.} 1997, in ASSL Vol. 210: The Impact of Large Scale
  Near-IR Sky Surveys, 25

\bibitem[{Smith {et~al.}(2002)Smith, Tucker, {et~al.}}]{smith02a}
Smith, J.~A., Tucker, D.~L., {et~al.} 2002, \aj, 123, 2121

\bibitem[{Stoughton {et~al.}(2002)}]{stoughton02a}
Stoughton, C. {et~al.} 2002, \aj, 123, 485

\bibitem[{Strauss {et~al.}(2002)}]{strauss02a}
Strauss, M.~A. {et~al.} 2002, \aj, 124, 1810

\bibitem[{{Witt} \& {Gordon}(2000)}]{witt00a}
{Witt}, A.~N. \& {Gordon}, K.~D. 2000, \apj, 528, 799

\bibitem[{York {et~al.}(2000)}]{york00a}
York, D. {et~al.} 2000, \aj, 120, 1579

\end{thebibliography}

@@

\subsection{Standard NMF}
The standard non-negative matrix factorization (NMF) problem, as
originally posed by \cite{lee00a}, is to approximate a data
matrix $\XX$ (of size $N_k\times N_n$) as the outer product of two
rank-$N_i$ non-negative matrices $\WW$ and $\HH$. For example, in this
paper, $\XX$ represents the fluxes at $N_n$ different wavelengths for
$N_k$ different galaxies, $\HH$ represents a set of $N_i$ templates to build
each galaxy from, and $\WW$ represents, for each galaxy, the weights
to give each template. This factorization is similar in spirit to
singular value decomposition (SVD) and principal components analysis
(PCA) but crucially, in NMF, $\XX$, $\WW$ and $\HH$ are all
constrained to have non-negative entries. This changes the problem
fundamentally, since no ``cancellation'' of positive and negative
basis functions or coefficients is possible -- all approximation
interactions are strictly additive. The goal, as with SVD or PCA, is
to minmize the squared approximation error (Frobenius norm):
\beqa \label{eq:nmfcost}
\chi^2 &= \|\XX-\WW\HH\|^2\\
%&= \|\XX-\XXh\|^2\\
&= \sum_{kn} \left(\Xkn - \sum_i \Wki\Hin \right)^2
\eeqa
In their seminal papers, \cite{lee00a} show that this approximation
error in non-increasing under the following very simple multiplicative
update rules: 
\beqa \label{eq:wupdate}
\Wki &\leftarrow \Wki \frac{[\XX\HH\T]_{ki}}{[\WW\HH\HH\T]_{ki}}\\
\label{eq:hupdate}
\Hin &\leftarrow \Hin \frac{[\WW\T\XX]_{in}}{[\WW\T\WW\HH]_{in}}
\eeqa
$[M]_{ab}$ denotes the $(ab)$ element of the matrix valued expression
$M$. $\HH\T$ is the transpose of matrix $\HH$.

@@

Our main algorithmic contribution is to derive a similar update equation
for the elements of $\BB$:
\beq \label{eq:bupdate}
\Bij \leftarrow \Bij
\frac{[\AA\T\XX\MM\T]_{ij}}{[\AA\T\AA\BB\MM\MM\T]_{ij}}
\eeq
Following the technique outlined in \cite{lee00a}, it can still be shown
that the error is nonincreasing under the application of this update.

The proof involves the use of an inequality lemma for symmetric
non-negative matrices:

{\bf Lemma:} 
{\it For any symmetric matrix $\PP$ having non-negative entries $\Pni
\geq 0$, any vector $\zz$ having non-negative entries $\zn \geq 0$,
and any vector $\yy$,}
\beq
\sum_{ni} \yn \yi \Pni \leq \sum_{ni} \yn^2 \frac{\zi}{\zn} \Pni 
\eeq
%{\it If the vector $\yy$ has non-negative entries, equality can be
%achieved by choosing $\zz=\yy$.}

For the standard NMF problem, we can prove that the approximation
error \eqref{eq:nmfcost} is non-increasing under \eqref{eq:wupdate} by
using the above lemma to construct a function $\phi(\WW,\ZZ)$ which is
an upper bound on the cost $\chi^2(\WW)$ for any non-negative matrix $\ZZ$:
{\small 
\beqa 
\chi^2(\WW) &= \sum_{kni\ell} \Wki\Wkl \Hin\Hln
- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
\phi(\WW,\ZZ) &= \sum_{kni\ell} \Wki^2\frac{\Zkl}{\Zki} \Hin\Hln 
- 2 \sum_{kni} \Wki \Xkn \Hin + \sum_{kn} \Xkn^2 \\
\phi(\WW,\ZZ) &\geq \chi^2(\WW) \qquad \forall \: \Wki\geq 0,\: \Zki\geq 0
\eeqa
} with equality being achieved when $\ZZ=\WW$. Figure \ref{lemma}
represents this definition schematically, showing the true $\chi^2$
function and $\phi$. The trick of the method is to define $\phi$ such
that it is equal to $\chi^2$ at $W$, greater than $\chi^2$ everywhere
else, and is easily minimizable. Then one can use the minimum of
$\phi$ as the update, and be guaranteed that one's updates do not
increase $\chi^2$.

The function $\phi$ can be analytically minimized with respect to
its first argument:
\beqa
\phi(\WW^*,\ZZ) &\leq \phi(\WW,\ZZ) \qquad \forall \: \WW,\ZZ \\
\Wki^* &= \Zki \frac{\sum_n \Xkn \Hin}{\sum_{n\ell} \Hin\Hln \Zkl}
\eeqa
If we set $\ZZ=\WW$, this becomes exactly \eqref{eq:wupdate} and
now we can easily prove the validity of this update rule: 
\beq
\chi^2(\WW) = \phi(\WW,\WW) \geq \phi(\WW^*,\WW) \geq \chi^2(\WW^*)
\eeq
where the first inequality comes from the fact that $\WW^*$ minimizes
$\phi(\cdot,\WW)$ with respect to its first argument
and the second comes from the fact that $\phi(\WW^*,\cdot)$ is a bound
on $\chi^2(\WW^*)$. The proof of validity for the update
\eqref{eq:hupdate} is analogous by symmetry.

To prove the validity of our update \eqref{eq:bupdate} for our new
problem, we proceed in a similar fashion, using the lemma twice to
construct consecutive upper bounds on the cost:
{\small 
\beqa 
\chi^2(\BB) &= \sum_{kni\ell{}js} \Aki\Akl\Bij\Bls\Mjn\Msn\\ \nonumber
&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
\phi(\BB,\ZZ,\beta) &= \sum_{kni\ell{}js} \Aki\Akl\Bij^2
\frac{\Zks\betalj}{\Zkj\betaij}\Mjn\Msn\\ \nonumber
&- 2 \sum_{knij} \Xkn\Aki\Bij\Mjn + \sum_{kn} \Xkn^2 \\
\phi(\BB,\ZZ,\beta) &\geq \chi^2(\BB) \qquad \forall 
\: \Bij\geq 0,\: \Zkj\geq 0,\: \betaij\geq 0
\eeqa
} with equality being achieved when $\ZZ=\AA\BB$ and 
$\beta=\BB$.

Once again, the bound $\phi$ can be analytically minimized with respect to
its first argument:
\beqa
\phi(\BB^*,\ZZ,\beta) &\leq \phi(\BB,\ZZ,\beta) \qquad 
\forall \: \BB,\ZZ,\beta \\
\Bij^* &= \betaij \frac{\sum_{kn} \Xkn\Aki\Mjn}
{\sum_{kn\ell{}s} \Aki\Akl\frac{\Zks\betalj}{\Zkj}\Mjn\Msn}
\eeqa
If we set $\ZZ=\AA\BB$ and $\beta=\BB$, this becomes
exactly \eqref{eq:bupdate} and 
we can now prove the validity of this update rule: 
\beq
\chi^2(\BB) = \phi(\BB,\AA\BB,\BB) \geq \phi(\BB^*,\AA\BB,\BB) 
\geq \chi^2(\BB^*) 
\eeq
where the first inequality comes from the fact that $\BB^*$ minimizes
$\phi(\cdot,\AA\BB,\BB)$ with respect to its first argument
and the second comes from the fact that $\phi(\BB^*,\cdot,\cdot)$ is a bound
on $\chi^2(\BB^*)$. The only condition we require is that the elements
of the matrix $\MM\MM\T$ are all non-negative.

@@

\clearpage
\begin{deluxetable}{rrrrr}
\tablewidth{0pt}
\tablecolumns{5}
\tablecaption{\label{solarmagnitudes} Properties of various filters}
\tablehead{ Band & $\lambda_{\mathrm{eff}}$ (\AA) & $m_{\mathrm{AB}} -
	m_{\mathrm{Vega}}$ & $M_\odot$ (AB) &
	$M_\odot$ (Vega)}
\startdata
$U$ & 3571. & 0.79 & 6.35 & 5.55\cr
$B$ & 4344. & -0.09 & 5.36 & 5.45\cr
$V$ & 5456. & 0.02 & 4.80 & 4.78\cr
$R$ & 6442. & 0.21 & 4.61 & 4.41\cr
$I$ & 7994. & 0.45 & 4.52 & 4.07\cr
$u$ & 3546. & 0.91 & 6.38 & 5.47\cr
$g$ & 4670. & -0.08 & 5.12 & 5.20\cr
$r$ & 6156. & 0.16 & 4.64 & 4.49\cr
$i$ & 7472. & 0.37 & 4.53 & 4.16\cr
$z$ & 8917. & 0.54 & 4.51 & 3.97\cr
$J$ & 12355. & 0.91 & 4.56 & 3.65\cr
$H$ & 16458. & 1.39 & 4.71 & 3.32\cr
$K_s$ & 21603. & 1.85 & 5.14 & 3.29\cr
\band{0.1}{u} & 3224. & 1.25 & 6.78 & 5.53\cr
\band{0.1}{g} & 4245. & -0.01 & 5.43 & 5.44\cr
\band{0.1}{r} & 5597. & 0.04 & 4.76 & 4.71\cr
\band{0.1}{i} & 6792. & 0.27 & 4.57 & 4.30\cr
\band{0.1}{z} & 8107. & 0.46 & 4.52 & 4.05\cr
\enddata
\tablecomments{Uses model solar spectrum and model Vega spectrum of
	\citet{kurucz91a}. Effective wavelength is defined in the text. The
	$UBRVI$ filters are those of \citet{bessell90a}. The $ugriz$ filters
	are those determined by Mamoru Doi, Daniel Eisenstein, and James
	Gunn and available on the SDSS DR4 web site.\footnote{{\tt
	http://www.sdss.org/dr4/}} The $JHK_s$ filters are those from
	\citet{cohen03a}. }
\end{deluxetable}

\begin{deluxetable}{lr}
\tablewidth{0pt}
\tablecolumns{2}
\tablecaption{\label{lineareqs} Conversions among various filters}
\tablehead{ Equation & Color dispersion }
\startdata
$u = \band{0.1}{u} - 0.3310 - 0.3014 \left[ (\band{0.1}{u}-\band{0.1}{g}) - 1.2839 \right] $ & $\sigma\left[\band{0.1}{u}-\band{0.1}{g}\right] = 0.28$ \cr
$g = \band{0.1}{g} - 0.3112 - 0.3530 \left[ (\band{0.1}{g}-\band{0.1}{r}) - 0.7187 \right] $ & $\sigma\left[\band{0.1}{g}-\band{0.1}{r}\right] = 0.20$ \cr
$r = \band{0.1}{r} - 0.2026 - 0.3810 \left[ (\band{0.1}{r}-\band{0.1}{i}) - 0.3530 \right] $ & $\sigma\left[\band{0.1}{r}-\band{0.1}{i}\right] = 0.07$ \cr
$i = \band{0.1}{i} - 0.1072 - 0.4273 \left[ (\band{0.1}{i}-\band{0.1}{z}) - 0.1914 \right] $ & $\sigma\left[\band{0.1}{i}-\band{0.1}{z}\right] = 0.13$ \cr
$z = \band{0.1}{i} - 0.3161 - 1.2057 \left[ (\band{0.1}{i}-\band{0.1}{z}) - 0.1914 \right] $ & $\sigma\left[\band{0.1}{i}-\band{0.1}{z}\right] = 0.13$ \cr
$u = U + 0.0682 + 0.0197 \left[ (U-B) - 0.9602 \right] $ & $\sigma\left[U-B\right] = 0.20$ \cr
$g = B - 0.2354 - 0.3411 \left[ (B-V) - 0.5870 \right] $ & $\sigma\left[B-V\right] = 0.18$ \cr
$r = V - 0.2585 - 0.5003 \left[ (V-R) - 0.3161 \right] $ & $\sigma\left[V-R\right] = 0.07$ \cr
$i = R - 0.2000 - 0.4248 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
$z = R - 0.4088 - 1.2495 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
$\band{0.1}{u} = U + 0.3989 + 0.4135 \left[ (U-B) - 0.9602 \right] $ & $\sigma\left[U-B\right] = 0.20$ \cr
$\band{0.1}{g} = U - 0.8845 - 0.9508 \left[ (U-B) - 0.9602 \right] $ & $\sigma\left[U-B\right] = 0.20$ \cr
$\band{0.1}{g} = B + 0.0759 + 0.0620 \left[ (B-V) - 0.5870 \right] $ & $\sigma\left[B-V\right] = 0.18$ \cr
$\band{0.1}{r} = B - 0.6429 - 1.0845 \left[ (B-V) - 0.5870 \right] $ & $\sigma\left[B-V\right] = 0.18$ \cr
$\band{0.1}{r} = V - 0.0558 - 0.1803 \left[ (V-R) - 0.3161 \right] $ & $\sigma\left[V-R\right] = 0.07$ \cr
$\band{0.1}{i} = R - 0.0927 + 0.0035 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
$\band{0.1}{z} = R - 0.2841 - 1.0301 \left[ (R-I) - 0.2652 \right] $ & $\sigma\left[R-I\right] = 0.12$ \cr
$\band{0.1}{u} = u + 0.3310 + 0.3203 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
$\band{0.1}{g} = u - 0.9528 - 0.7572 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
$\band{0.1}{g} = g + 0.3113 + 0.4620 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
$\band{0.1}{r} = g - 0.4075 - 0.8577 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
$\band{0.1}{i} = r - 0.1504 - 0.3654 \left[ (r-i) - 0.2589 \right] $ & $\sigma\left[r-i\right] = 0.10$ \cr
$\band{0.1}{z} = i - 0.0836 - 0.7518 \left[ (i-z) - 0.2083 \right] $ & $\sigma\left[i-z\right] = 0.10$ \cr
$U = u - 0.0682 - 0.0140 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
$B = u - 1.0286 - 0.7981 \left[ (u-g) - 1.2638 \right] $ & $\sigma\left[u-g\right] = 0.26$ \cr
$B = g + 0.2354 + 0.3915 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
$V = g - 0.3516 - 0.7585 \left[ (g-r) - 0.6102 \right] $ & $\sigma\left[g-r\right] = 0.15$ \cr
$R = r - 0.0576 - 0.3718 \left[ (r-i) - 0.2589 \right] $ & $\sigma\left[r-i\right] = 0.10$ \cr
$I = i - 0.0647 - 0.7177 \left[ (i-z) - 0.2083 \right] $ & $\sigma\left[i-z\right] = 0.10$ \cr
$U = \band{0.1}{u} - 0.3992 - 0.3189 \left[ (\band{0.1}{u}-\band{0.1}{g}) - 1.2839 \right] $ & $\sigma\left[\band{0.1}{u}-\band{0.1}{g}\right] = 0.28$ \cr
$B = \band{0.1}{g} - 0.0759 - 0.0545 \left[ (\band{0.1}{g}-\band{0.1}{r}) - 0.7187 \right] $ & $\sigma\left[\band{0.1}{g}-\band{0.1}{r}\right] = 0.20$ \cr
$V = \band{0.1}{g} - 0.6628 - 0.9259 \left[ (\band{0.1}{g}-\band{0.1}{r}) - 0.7187 \right] $ & $\sigma\left[\band{0.1}{g}-\band{0.1}{r}\right] = 0.20$ \cr
$R = \band{0.1}{r} - 0.2603 - 0.9162 \left[ (\band{0.1}{r}-\band{0.1}{i}) - 0.3530 \right] $ & $\sigma\left[\band{0.1}{r}-\band{0.1}{i}\right] = 0.07$ \cr
$I = \band{0.1}{i} - 0.1725 - 0.9718 \left[ (\band{0.1}{i}-\band{0.1}{z}) - 0.1914 \right] $ & $\sigma\left[\band{0.1}{i}-\band{0.1}{z}\right] = 0.13$ \cr
\enddata
\tablecomments{Uses AB magnitudes throughout.}
\end{deluxetable}
